Vertical shear instability in astrophysical disks: vertically global linear calculations with thermal relaxation and applications to the Solar Nebula

### 2015-03-03 linear_iso.tex
updating main part of paper
almost done with version 2

\section{Analytical discussion of vertically isothermal
  disks in the framework of the vertically global shearing
  box}\label{analytical}  
We begin with a discussion of vertically isothermal disks
where $\Gamma=1$, then $c_s$ is a constant in the radially-local
approximation. We  first examine isothermal perturbations ($\beta =
0$) before considering the effect of finite thermal relaxation. 
In order to facilitate a fully analytic discussion, we introduce a
number of approximations below that allows a simple governing equation
to be derived and compared with previous works.

\subsection{Relation to the vertically global
  shearing box} 
We connect the above linearized equations to the vertically global
shearing box (VGSB) formalism developed by \cite{mcnally14}. The VGSB
is an extension of the standard shearing box \citep{goldreich65} to
allow the shear flow in the shearing box to be height-dependent.  
The difference between the radially-local approximation and the VGSB is
that the radial disk structure is ignored in the latter.  That is, the
equilibrium VGSB is homogeneous in the radial direction, as in the
standard shearing box.

Thus, the linear problem in the VGSB can be derived from
Eq. \ref{lin_mass}---\ref{lin_energy} by setting 
$\hat{g}_c=0$.  
The resulting linear problem is very similar
to that for local axisymmetric waves in accretion disks already
discussed by \cite{lubow93}, but modified by the presence of vertical
shear and thermal relaxation. This is convenient as it allows us to
directly apply their solution methods.

@@

The VGSB is a self-consistent framework but extrapolating results 
obtained from the VGSB to the global disk has limitations. This is
because within the VGSB framework, $d\Omega/dz$ can be freely
specified. In effect, by adopting the VGSB framework, we are studying
dynamics in the standard shearing box subject to an additional
azimuthal force $\delta v_zr d \Omega/dz$ with $d\Omega/dz$ as an
input parameter. This is different to the global problem because in
that case, the vertical shear is a result of the background radial
disk structure, which is ignored in the VGSB.

@@

\subsection{Governing equation in the VGSB}\label{approx_gov}
We now derive a governing equation with further approximations:
\begin{enumerate}
\item We consider eigenvalues such that
  $|\sigma^2|\ll \kappa^2$. Then we can replace $D=\kappa^2 -\sigma^2\to
  \kappa^2$. % The problem simplifies because the eigenvalue now only
  % appears in the last term in Eq. \ref{iso_ode}.
  We call this the
  low-frequency approximation. 
\item We are mainly interested in the effect of vertical shear, i.e. $q\neq
  0$. This is already captured explicitly in the linearized equations through
  the term $d\Omega/dz$. We therefore simplify
  further by ignoring the vertical dependence of $\Omega$ and $\kappa$
  elsewhere. For clarity, we replace $\kappa^2$ with $\Omega_k^2$.

@@

It is convenient to introduce the dimensionless variables
\begin{align}
  \hat{z} = z/\Hiso,\quad \hat{k}=k_x\Hiso, \quad \hat{\sigma} = \sigma/\Omega_k,
\end{align}
then Eq. \ref{vertiso_gov} becomes
\begin{align}
  0=& \delta v_z ^{\prime\prime} + \left(1 + \ii\epsilon
    q\hat{k}\right)\ln\rho^{\prime}\delta v_z^\prime \notag\\
  &+
  \left[\hat{\sigma}^2\left(\hat{k}^2+\chi\right) +
    \left(  \chi + \ii \epsilon q\hat{k}\right)\ln\rho^{\prime\prime}\right.
  \notag\\
  &\left.- \ln\rho^{\prime
      2}\left(\khat^2 -
      \ii\epsilon
      q\hat{k}\right)\left(1 - \chi\right)\right]\delta v_z,\label{vertiso_gov_nondim}
\end{align}
where $^\prime$ denotes $d/d\zhat$. Eq. \ref{vertiso_gov_nondim} will
be the basis for our discussion of vertically isothermal disks. % ,
% although further simplifications or maniputations will be performed
% depending the type of perturbations being considered.

In deriving Eq. \ref{vertiso_gov}, the vertical dependence of
$D=\kappa^2(z)-\sigma^2$ was ignored because the replacement
$D\to\Omega_k^2$ was made before eliminating variables in favor of
$\delta v_z$. In Appendix \ref{adia_improve} we show that making this
replacement after eliminating variables, which requires the vertical
derivative of $D$, adds unnecessary complexity for thin disks.

@@

Note that we can generalize Eq. \ref{vertiso_gov_nondim} to other
(non-Keplerian) rotation profiles by redefining $\khat$
as $k_xc_s/\kappa$, $\hat{\sigma}$ as $\sigma/\kappa$, and 
$\epsilon$ as $c_s/r\kappa$.

@@

\subsection{Isothermal perturbations}\label{iso_discuss}
Isothermal perturbations correspond to $\beta\equiv0$, 
%relaxation timescale $\be
%or for any adiabatc index $\gamma$ when
%$\beta=0$. 
so $\chi=1$ and Eq. \ref{ode_Q} implies 
\begin{align}
  Q = W, 
\end{align}
 and so 
 \begin{align}
   \ii\hat{\sigma}c_s\delta v_z = W^\prime
   % \quad   \text{(isothermal perturbations)},
 \end{align}
from Eq. \ref{lin_vz}. For isothermal perturbations it is in fact 
simpler to work with $W$ by integrating Eq. \ref{vertiso_gov_nondim}. For later discussion
it is convenient to write the result as

\begin{align}
  0 = W^{\prime\prime} + \left[\ln\rho^\prime - \ii \epsilon q \hat{k}
    f(\zhat)\right]W^\prime + \hat{\sigma}^2\left(1+\hat{k}^2\right)W,\label{iso_ode1}
\end{align}
where $f(\zhat)$ is defined such that
\begin{align}\label{fz_shear}
  \frac{d\Omega^2}{d\hat{z}} = \epsilon^2q f(\hat{z})\Omega_k^2.
\end{align}

% The linear problem simplifies considerably for the case
% $\Gamma=\gamma=1$.  Then 
% \begin{align}
%   Q=W. 
% \end{align}
% Inserting this into Eq. \ref{ode_w} gives a single equation for the linear stability
% problem for isothermal perturbations in a vertically isothermal disk, 
% \begin{align}\label{iso_ode}
%   \frac{d^2W}{dz^2} + \left(\frac{d\ln{\rho}}{dz} - \frac{\ii k_x
%       r}{D}\frac{d\Omega^2}{dz}\right) \frac{dW}{dz} +
%   \sigma^2\left(\frac{1}{c_s^2} + \frac{k_x^2}{D}\right)W=0, 
% \end{align} 
% subject to appropriate boundary conditions. Eq. \ref{iso_ode} also
% applies to $\gamma>1$ for $t_c\to0$.

Note that for isothermal perturbations the issue of ignoring the
vertical dependence of $D$ described in \S\ref{approx_gov} and Appendix
\ref{adia_improve} is irrelevant, since
Eq. \ref{iso_ode1} may be derived directly by substituting $Q=W$ in
Eq. \ref{ode_w}, i.e. no derivatives of $D$ needs to be taken.

@@

\subsubsection{Maximum growth rate}   
Here we show that the growth rate is limited by the maximum vertical
shear in the domain. Writing $\hat{\sigma} = \hat{\omega} +
\ii\hat{\nu}$ with real $\hat{\omega}$ and $\hat{\nu}$, the real and
imaginary parts of 
Eq. \ref{integral_relation1} are
\begin{align}
  &\left(\hat{\omega}^2-\hat{\nu}^2\right)\left(1+\hat{k}^2\right)
  \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z} -
  \int_{\zhat_1}^{\zhat_2}\rho|W^\prime|^2d\hat{z}
  \notag\\
  &=\real\left[\ii
    \epsilon\hat{k}q \int_{\zhat_1}^{\zhat_2}\rho
    f(\hat{z}) W^*W^\prime d\zhat\right], \notag\\
  & 2\hat{\omega}\hat{\nu}\left(1+\hat{k}^2\right)
  \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z}\notag\\
  &=\imag\left[\ii
    \epsilon\hat{k}q \int_{\zhat_1}^{\zhat_2}\rho
    f(\hat{z}) W^*W^\prime d\hat{z}\right].
\end{align}
Adding the square of these equations give
\begin{align}
  &\left[|\hat{\sigma}|^2\left(1+\hat{k}^2\right)
    \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z} -
    \int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}\right]^2\notag\\
  &+4\hat{\nu}^2\left(1+\hat{k}^2\right) 
  \int_{\zhat_1}^{\zhat_2}\rho
  |W|^2d\hat{z}\int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}\notag\\
  &=\left|\ii
    \epsilon\hat{k}q\int_{\zhat_1}^{\zhat_2}\rho
    f(\hat{z}) W^*W^\prime d\hat{z}\right|^2.
\end{align}
It is clear that
\begin{align}\label{sigma_finite_domain} 
  &4\hat{\nu}^2\left(1+\hat{k}^2\right) 
  \int_{\zhat_1}^{\zhat_2}\rho
  |W|^2d\hat{z}\int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}\notag\\
  &\leq\left|
    \epsilon\hat{k}q\int_{\zhat_1}^{\zhat_2}\rho
    f(\hat{z}) W^*W^\prime d\hat{z}\right|^2.
\end{align}
On the left hand side of this inequality, we apply the Cauchy-Schwarz
inequality to obtain
\begin{align}
  &\left( \int_{\zhat_1}^{\zhat_2}\rho
    |W|\left|W^\prime \right|d\hat{z}\right)^2\leq
  \int_{\zhat_1}^{\zhat_2}\rho 
  |W|^2d\hat{z}\int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}.
\end{align}
On the right hand side of Eq. \ref{sigma_finite_domain} we have
\begin{align}
  \left|\int_{\zhat_1}^{\zhat_2}\rho
    f(\hat{z}) W^*W^\prime d\hat{z}\right|\leq \int_{\zhat_1}^{\zhat_2}\rho
  \left|f(\hat{z})W^*W^\prime \right|d\hat{z} \notag\\
  \leq
  \mathrm{max}\left(|f|\right)\int_{\zhat_1}^{\zhat_2}\rho
  |W|\left|W^\prime \right|d\hat{z},
\end{align}
where $\mathrm{max}(|f|)$ is the maximum value of $|f|$ in
$\zhat\in[\zhat_1,\zhat_2]$. Inserting these inequalities into
Eq. \ref{sigma_finite_domain} gives
\begin{align}\label{max_growth}
  |\hat{\nu}|\leq
  \frac{\epsilon |\hat{k} q|}{2\sqrt{1+\hat{k}^2}}\mathrm{max}(|f|). 
\end{align}
Recalling that $f$ represents vertical shear (Eq. \ref{fz_shear}), it
follows that the maximum possible growth rate of unstable modes,
satisfying the above boundary conditions, is limited by the maximum
value of the vertical shear in the domain considered.

For the vertically isothermal disk we have $|f(\hat{z})| =
|\hat{z}|\left(1+\epsilon^2\hat{z}^2\right)^{-3/2}$, which maximizes
at $|\zhat|=1/\sqrt{2}\epsilon$. For vertical domains smaller 
than this height, the growth rate is limited by $|f|$ at the vertical
boundaries considered. For $|\epsilon\hat{z}|\ll1$ we have $f\simeq
\hat{z}$, in which case we can expect growth rates to increase at most linearly
with respect to the vertical domain size.   %no faster than linear

@@

\subsubsection{Stability in the absence of vertical shear}\label{iso_stable}
When $q=0$, Eq. \ref{iso_ode3} is
\begin{align}\label{hermite_ode}
  \left[w(\hat{z})W^\prime \right]^\prime + nW
  w(\hat{z}) =0, 
\end{align}
where we have defined a new eigenvalue
\begin{align}
  n \equiv \hat{\sigma}^2(1+\hat{k}^2). 
\end{align} 
Eq. \ref{hermite_ode} is Hermite's differential equation. If we impose
that the kinetic energy density remain bounded at infinity, then  
\begin{align}
  W \propto \He_n(\hat{z}),
\end{align}
and $n$ is a non-negative integer. This translates to a real
eigenfrequency $\sigma$ such that
\begin{align}
  \left|\hat{\sigma}\right| = \sqrt{n}
  \left(1+\hat{k}^2\right)^{-1/2}. \label{iso_stable_disp}
\end{align}
Since we have assumed $|\sigma^2|\ll \kappa^2\sim \Omega_k^2$, our
analysis is only valid for large wavenumbers such that $\hat{k}^2\gg   
n-1$. Physically, this corresponds to radial length-scales much
smaller than the local disk scale height.

We remark that for fixed $n$ and $\khat\gg 1$ the dispersion relation 
Eq. \ref{iso_stable_disp} imply $\hat{\omega}\propto 1/\khat$. This inverse relation 
has also been qualitatively observed in numerical simulations performed by 
\cite{stoll14}.

\subsubsection{Polynomial solutions with vertical shear}\label{iso_poly}
We can seek power-series solution to Eq. \ref{iso_ode3},
\begin{align}
  W(\zhat) = \sum_{l=0}^\infty a_l\zhat^l. 
\end{align}
Then the coefficients must satisfy the recurrence relation
\begin{align}
  (l+2)(l+1)a_{l+2} +
  \left[n - l\left(1+\ii \epsilon q  \hat{k}\right)\right] a_l = 0, 
\end{align}
where $n$ is not necessarily an integer. Indeed, if we demand
a polynomial of degree $L$ as the solution, then the eigenfrequency
must satisfy
\begin{align}\label{sig2_iso}
\hat{\sigma}^2 = L\left(\frac{1+\ii q \epsilon
    \hat{k}}{1+\hat{k}^2}\right).
\end{align}
The eigenfrequency $\hat{\sigma}$ is therefore complex for
$L\geq1$. We are interested in 
unstable modes for which $\imag{\hat{\sigma}}>0$ is the growth rate and is
given by 
\begin{align}\label{simple_growth}
  \hat{\nu} =\sqrt{
   \frac{L}{2\left(1+\hat{k}^2\right)}\left(\sqrt{1+q^2\epsilon^2\hat{k}^2} - 
    1\right)}. 
\end{align}
For fixed $q$ and $\epsilon$, the growth rate vanishes for both
$\hat{k}^2\to0$ and $\hat{k}^2\to\infty$. The maximum growth rate
occurs at the optimum wavenumber $\hat{k}_\mathrm{opt}$
\begin{align}
  |\hat{k}_\mathrm{opt}| = \sqrt{\frac{2+|\epsilon q|}{|\epsilon q|}},
\end{align}
and
\begin{align}
  \mathrm{max}\left(\hat{\nu}\right) =\frac{\sqrt{L}|\epsilon
    q|}{2\sqrt{1+|\epsilon q|}}. \label{iso_max_growth}
\end{align}
Since $L\geq1$, there is no limit to the maximum growth rate. This is
an artifact of the thin-disk approximation because the vertical 
shear $d\Omega^2/dz\propto z$ increases indefinitely with
height. However, we find Eq. \ref{iso_max_growth} compares well with
numerical solutions of the governing equation for small $L$.

>>>>

\section{Analytical discussion of vertically isothermal
  disks in the framework of the vertically global shearing
  box}\label{analytical}  
We begin with a discussion of vertically isothermal disks
where $\Gamma=1$, then $c_s$ is a constant in the radially-local
approximation. We  first examine isothermal perturbations ($\beta =
0$) before considering the effect of finite thermal relaxation. 
In order to facilitate a fully analytic discussion, we first introduce a
number of approximations below that allows a simple governing equation
to be derived and compared with previous works.

\subsection{Relation to the vertically global
  shearing box} 
We connect the linearized equations 
Eq. \ref{lin_mass}---\ref{lin_energy} to the vertically global  
shearing box (VGSB) formalism developed by \cite{mcnally14}. The VGSB
is an extension of the standard shearing box \citep{goldreich65} to
allow the shear flow in the shearing box to be height-dependent.  
The difference between the radially-local approximation and the VGSB is
that the radial disk structure is ignored in the latter. That is, the 
equilibrium density and pressure field in the VGSB is homogeneous in
the radial ($x$) direction, as in the standard shearing box.

Thus, the linear problem in the VGSB can be derived from
Eq. \ref{lin_mass}---\ref{lin_energy} by setting 
$\hat{g}_c=0$. The resulting linear problem is very similar 
to that for local axisymmetric waves in accretion disks already
discussed by \cite{lubow93}, but modified by the presence of vertical
shear and thermal relaxation. This is convenient as it allows us to
directly apply some of their solution methods.

@@

The VGSB is a self-consistent framework but extrapolating results 
obtained from the VGSB to the global disk has limitations. This is
because within the VGSB framework, $d\Omega/dz$ can be freely
specified. In effect, by adopting the VGSB framework, we are studying
dynamics in the standard shearing box subject to an additional
azimuthal force $\delta v_zr d \Omega/dz$ with $d\Omega/dz$ as an
input parameter. For consistency with the global disk, we choose
$d\Omega/dz$ in the VGSB to be that due to the thermal structure in
the global disk (Eq. \ref{vertical_shear}). However, 
% This is different to the global problem because in
% that case, the vertical shear is a result of the 
the background radial disk structure is ignored elsewhere in the VGSB.

@@

\subsection{Governing equation in the VGSB}\label{approx_gov}
We now derive a governing equation with further approximations:
\begin{enumerate}
\item We consider eigenvalues such that
  $|\sigma^2|\ll \kappa^2$. Then we can replace $D=\kappa^2 -\sigma^2\to
  \kappa^2$. % The problem simplifies because the eigenvalue now only
  % appears in the last term in Eq. \ref{iso_ode}. %stoll kley see low
  % freq 
  We call this the low-frequency approximation. 
\item We are mainly interested in the effect of vertical shear, i.e. $q\neq
  0$. This is already captured explicitly in the linearized equations through
  the term $d\Omega/dz$. We therefore simplify 
  further by ignoring the vertical dependence of $\Omega$ and $\kappa$
  elsewhere. For clarity, we replace $\kappa^2$ with $\Omega_k^2$.

@@

It is convenient to introduce the dimensionless variables
\begin{align}
  \hat{z} = z/H,\quad \hat{k}=k_xH, \quad \hat{\sigma} = \sigma/\Omega_k,
\end{align}
then Eq. \ref{vertiso_gov} becomes
\begin{align}
  0=& \delta v_z ^{\prime\prime} + \left(1 + \ii\epsilon
    q\hat{k}\right)\ln\rho^{\prime}\delta v_z^\prime \notag\\
  &+
  \left[\hat{\sigma}^2\left(\hat{k}^2+\chi\right) +
    \left(  \chi + \ii \epsilon q\hat{k}\right)\ln\rho^{\prime\prime}\right.
  \notag\\
  &\left.- \ln\rho^{\prime
      2}\left(\khat^2 -
      \ii\epsilon
      q\hat{k}\right)\left(1 - \chi\right)\right]\delta v_z,\label{vertiso_gov_nondim}
\end{align}
where $^\prime$ denotes $d/d\zhat$. Eq. \ref{vertiso_gov_nondim} will
be the basis for our discussion of vertically isothermal disks. % ,
% although further simplifications or maniputations will be performed
% depending the type of perturbations being considered.

In deriving Eq. \ref{vertiso_gov}, the vertical dependence of
$D=\kappa^2(z)-\sigma^2$ was ignored because the replacement
$D\to\Omega_k^2$ was made before eliminating variables in favor of
$\delta v_z$. In Appendix \ref{adia_improve} we show that making this
replacement after eliminating variables, which involves $dD/dz$,
introduces terms $O(\epsilon^2)$. Since we are interested in 
thin disks ($\epsilon\ll 1 $), e.g. protoplaneary disks where 
$\epsilon\lesssim 0.1$ \citep{chiang10}, these terms introduce
unnecessary complexity for the present discussion. However, this
approximation will be relaxed in our numerical study.

@@

Note that we can generalize Eq. \ref{vertiso_gov_nondim} non-Keplerian
rotation profiles by redefining $\khat$ as $k_xc_s/\kappa$,
$\hat{\sigma}$ as $\sigma/\kappa$, and  $\epsilon$ as $c_s/r\kappa$.

@@

\subsection{Isothermal perturbations}\label{iso_discuss}
Isothermal perturbations correspond to $\beta\equiv0$ 
%relaxation timescale $\be
%or for any adiabatc index $\gamma$ when
%$\beta=0$. 
($\chi=1$) and Eq. \ref{ode_Q} becomes  
\begin{align}
  Q = W 
\end{align}
(i.e. $\delta P = c_s^2\delta \rho$), and so  
\begin{align}
  \ii\hat{\sigma}c_s\delta v_z = W^\prime
  % \quad   \text{(isothermal perturbations)},
\end{align}
from Eq. \ref{lin_vz}. For isothermal perturbations it is in fact 
simpler to work with $W$ by integrating Eq. \ref{vertiso_gov_nondim}. For later discussion
it is convenient to write the result as

\begin{align}
  0 = W^{\prime\prime} + \left[\ln\rho^\prime - \ii \epsilon q \hat{k}
    f(\zhat)\right]W^\prime + \hat{\sigma}^2\left(1+\hat{k}^2\right)W,\label{iso_ode1}
\end{align}
where $f(\zhat)$ is defined such that
\begin{align}\label{fz_shear}
  \frac{d\Omega^2}{d\hat{z}} = \epsilon^2q f(\hat{z})\Omega_k^2.
\end{align}
For a vertically isothermal disk we have 
\begin{align}
  f(\zhat) = \frac{\zhat}{\left(1+\epsilon^2\hat{z}^2\right)^{3/2}}, 
\end{align}
%For the vertically isothermal disk we have $|f(\hat{z})| =
%|\hat{z}|\left(1+\epsilon^2\hat{z}^2\right)^{-3/2}$, 
which maximizes at $|\zhat|=1/\sqrt{2}\epsilon$. The existence of a
maximum vertical shear limits the growth rate of unstable modes, as  
shown below.

Note that Eq. \ref{iso_ode1} can in fact be derived by substituting
$Q=W$ in Eq. \ref{ode_w} without approximating $D$ or $\kappa^2$ as
vertically constant.

% the issue of ignoring the
% vertical dependence of $D$ described in \S\ref{approx_gov} and Appendix
% \ref{adia_improve} is irrelevant, since
% Eq. \ref{iso_ode1} may be derived directly , i.e. no derivatives of $D$ needs to be taken.

@@

\subsubsection{Maximum growth rate}   
Here we show that the growth rate is limited by the maximum vertical
shear in the domain. Writing $\hat{\sigma} = \hat{\omega} +
\ii\hat{\nu}$ with real $\hat{\omega}$ and $\hat{\nu}$, the real and
imaginary parts of 
Eq. \ref{integral_relation1} are
\begin{align}
  &\left(\hat{\omega}^2-\hat{\nu}^2\right)\left(1+\hat{k}^2\right)
  \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z} -
  \int_{\zhat_1}^{\zhat_2}\rho|W^\prime|^2d\hat{z}
  \notag\\
  &=\real\left[\ii
    \epsilon\hat{k}q \int_{\zhat_1}^{\zhat_2}\rho
    f(\hat{z}) W^*W^\prime d\zhat\right], \notag\\
  & 2\hat{\omega}\hat{\nu}\left(1+\hat{k}^2\right)
  \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z}\notag\\
  &=\imag\left[\ii
    \epsilon\hat{k}q \int_{\zhat_1}^{\zhat_2}\rho
    f(\hat{z}) W^*W^\prime d\hat{z}\right].
\end{align}
Adding the square of these equations give
\begin{align}
  &\left[|\hat{\sigma}|^2\left(1+\hat{k}^2\right)
    \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z} -
    \int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}\right]^2\notag\\
  &+4\hat{\nu}^2\left(1+\hat{k}^2\right) 
  \int_{\zhat_1}^{\zhat_2}\rho
  |W|^2d\hat{z}\int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}\notag\\
  &=\left|\ii
    \epsilon\hat{k}q\int_{\zhat_1}^{\zhat_2}\rho
    f(\hat{z}) W^*W^\prime d\hat{z}\right|^2.
\end{align}
It is clear that
\begin{align}\label{sigma_finite_domain} 
  &4\hat{\nu}^2\left(1+\hat{k}^2\right) 
  \int_{\zhat_1}^{\zhat_2}\rho
  |W|^2d\hat{z}\int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}\notag\\
  &\leq\left|
    \epsilon\hat{k}q\int_{\zhat_1}^{\zhat_2}\rho
    f(\hat{z}) W^*W^\prime d\hat{z}\right|^2.
\end{align}
On the left hand side of this inequality, we apply the Cauchy-Schwarz
inequality to obtain
\begin{align}
  &\left( \int_{\zhat_1}^{\zhat_2}\rho
    |W|\left|W^\prime \right|d\hat{z}\right)^2\leq
  \int_{\zhat_1}^{\zhat_2}\rho 
  |W|^2d\hat{z}\int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}.
\end{align}
On the right hand side of Eq. \ref{sigma_finite_domain} we have
\begin{align}
  \left|\int_{\zhat_1}^{\zhat_2}\rho
    f(\hat{z}) W^*W^\prime d\hat{z}\right|\leq \int_{\zhat_1}^{\zhat_2}\rho
  \left|f(\hat{z})W^*W^\prime \right|d\hat{z} \notag\\
  \leq
  \mathrm{max}\left(|f|\right)\int_{\zhat_1}^{\zhat_2}\rho
  |W|\left|W^\prime \right|d\hat{z},
\end{align}
where $\mathrm{max}(|f|)$ is the maximum value of $|f|$ in
$\zhat\in[\zhat_1,\zhat_2]$. Inserting these inequalities into
Eq. \ref{sigma_finite_domain} gives
\begin{align}\label{max_growth}
  |\hat{\nu}|\leq
  \frac{\epsilon |\hat{k} q|}{2\sqrt{1+\hat{k}^2}}\mathrm{max}(|f|). 
\end{align}
Recalling that $f$ represents vertical shear (Eq. \ref{fz_shear}), it
follows that the maximum possible growth rate of unstable modes,
satisfying the above boundary conditions, is limited by the maximum
vertical shear in the domain considered.

In practice, one might consider a vertical domain of a few scale 
heights in a thin disk with $|q|=O(1)$. In this case, $|\epsilon\zhat|\ll1$ and  
$f\simeq \zhat$, so that $\mathrm{max}|f| = O(1)$, implying a
maximum growth rate $O(\epsilon \Omega_k)$, consistent with numerical
results below.

% For vertical domains smaller 
% than this height, the growth rate is limited by $|f|$ at the vertical
% boundaries considered. For $|\epsilon\hat{z}|\ll1$ we have $f\simeq
% \hat{z}$, in which case we can expect growth rates to increase at most linearly
% with respect to the vertical domain size.   %no faster than linear

@@

\subsubsection{Stability in the absence of vertical shear}\label{iso_stable}
When $q=0$, Eq. \ref{iso_ode3} is
\begin{align}\label{hermite_ode}
  \left[w(\hat{z})W^\prime \right]^\prime + nW
  w(\hat{z}) =0, 
\end{align}
where we have defined a new eigenvalue
\begin{align}
  n \equiv \hat{\sigma}^2(1+\hat{k}^2). 
\end{align} 
Eq. \ref{hermite_ode} is Hermite's differential equation. If we impose
that the kinetic energy density remain bounded at infinity, then the
solutions are Hermite polynomials  
\begin{align}
  W \propto \He_n(\hat{z}),
\end{align}
and $n$ is a non-negative integer. This translates to a real
eigenfrequency $\sigma$ such that
\begin{align}
  \left|\hat{\sigma}\right| = \sqrt{n}
  \left(1+\hat{k}^2\right)^{-1/2}. \label{iso_stable_disp}
\end{align}
Since we have assumed $|\sigma^2|\ll \kappa^2\sim \Omega_k^2$, our
analysis is only valid for large wavenumbers such that $\hat{k}^2\gg   
n-1$. Physically, this corresponds to radial length-scales much
smaller than the local disk scale height. Note that
Eq. \ref{iso_stable_disp} is implied by Eq. 54 in
\cite{lubow93} after setting $\gamma=1$ and making the low-frequency,
Keplerian disk approximation in that equation.

We remark that for fixed $n$ and $\khat\gg 1$ the dispersion relation 
Eq. \ref{iso_stable_disp} imply $\hat{\omega}\propto 1/\khat$. This inverse relation 
has also been qualitatively observed in numerical simulations
\citep{stoll14}.

\subsubsection{Polynomial solutions with vertical shear}\label{iso_poly}
We can seek power-series solution to Eq. \ref{iso_ode3},
\begin{align}
  W(\zhat) = \sum_{l=0}^\infty a_l\zhat^l. 
\end{align}
Then the coefficients must satisfy the recurrence relation
\begin{align}
  (l+2)(l+1)a_{l+2} +
  \left[n - l\left(1+\ii \epsilon q  \hat{k}\right)\right] a_l = 0, 
\end{align}
where $n$ is not necessarily an integer. Indeed, if we demand
a polynomial of degree $L$ as the solution, then the eigenfrequency
must satisfy
\begin{align}\label{sig2_iso}
\hat{\sigma}^2 = L\left(\frac{1+\ii q \epsilon
    \hat{k}}{1+\hat{k}^2}\right).
\end{align}
The eigenfrequency $\hat{\sigma}$ is therefore complex for
$L\geq1$. We are interested in 
unstable modes for which $\hat{\nu}=\imag{\hat{\sigma}}>0$ is the
growth rate and is given by 
\begin{align}\label{simple_growth}
  \hat{\nu} =\sqrt{
   \frac{L}{2\left(1+\hat{k}^2\right)}\left(\sqrt{1+q^2\epsilon^2\hat{k}^2} - 
    1\right)}. 
\end{align}
For fixed $q$ and $\epsilon$, the growth rate vanishes for both
$\hat{k}^2\to0$ and $\hat{k}^2\to\infty$. 
The maximum growth rate
occurs at the optimum wavenumber $\hat{k}_\mathrm{opt}$
\begin{align}
  |\hat{k}_\mathrm{opt}| = \sqrt{\frac{2+|\epsilon q|}{|\epsilon q|}},
\end{align}
and
\begin{align}
  \mathrm{max}\left(\hat{\nu}\right) =\frac{\sqrt{L}|\epsilon
    q|}{2\sqrt{1+|\epsilon q|}}. \label{iso_max_growth}
\end{align}
Since $L\geq1$, there is no limit to the maximum growth rate. This is
an artifact of the thin-disk approximation because the vertical 
shear $d\Omega^2/dz\propto z$ increases indefinitely with
height. However, we find Eq. \ref{iso_max_growth} compares well with
numerical solutions of the governing equation for small $L$.
